Theory about Normal Basis on $F(2^k)$

### 2017-01-24 dissertation/submission/prelim_GF.tex
dissertation: after submission_1st revision

Fields of the form $\F_{p^{k}}$ are called Galois extension fields.
We are specifically interested in extension fields of type 
$\Fkk$, where $k > 1$. These are extensions of the binary
field $\F_2$.

\begin{table}[!h]
\caption{Addition and multiplication operations over $\F_2$.}
	\centering
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $1$ & \\
	& $1$ & $1$ & $0$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Addition over $\F_2$}\\
	\end{tabular}
	\quad
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $0$ & \\
	& $1$ & $0$ & $1$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
	\end{tabular}
\end{table}

Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
because it is performed modulo $2$.
Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.

@@

\begin{table}[h]
\begin{center}
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$.}\label{tab:gfelement} 
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
\caption{A typical SMSO structure of Massey-Omura multiplier.}
\label{fig:MasseyOmura}}
\end{figure}

@@

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
\caption{5-bit Agnew's SMPO. Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$. Index $w = 2i\pmod 5$.}
\label{fig:SMPO}}
\end{figure}

@@

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
\caption{A 5-bit RH-SMPO.}
\label{fig:5bitRH}}
\end{figure}

@@

\begin{algorithm}[H]
\SetAlgoNoLine

 \KwIn{$A,B\in \Fkk$ given {\it w.r.t.} NB $N$}
 \KwOut{$R=A\times B$}
%%%%%%%%%%%%%%%%%%%%
  Initialize $A,B$ and aux var $X$ to 0\;
  \For { ($i=0$;   $i < k$; ++i ) }
  {
	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Equation \ref{eqn:aux}*/\;}
	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
  }
  $R\gets X$
\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
\end{algorithm}

In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
This is because of equation
$$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
$F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
\end{Proof}

To better understand the mechanism of RH-SMPO, we will use this 5-bit 
RH-SMPO as an example and introduce the details on how to design it.
\begin{Example}[Designing a 5-bit RH-SMPO]
From Equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
and XOR gates in $c$-layer according to Equation \ref{eqn:auxDC}. Concretely, as Algorithm \ref{alg:RHmulti}
describes, we implement auxiliary function $F_{k-1}$ in the logic:
$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
\begin{equation}
\label{eqn:5bitRHaux}
F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
\end{equation}
Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
$i$-th gate forms the coefficient of $\beta^{2^i}$.

@@

\begin{figure}[H]
\centering{
\includegraphics[width=3in]{newfig/multitable.pdf}
\caption{A $5\times 5$ multiplication table.}
\label{fig:multitable}}
\end{figure}

>>>>

Fields of the form $\F_{p^{k}}$ are called Galois extension fields.
We are specifically interested in extension fields of type 
$\Fkk$, where $k > 1$, because they are extensions of the binary
field $\F_2$. Elements and operators in $\F_2$ can be mapped to 
Boolean values and Boolean operators, as Table \ref{tab:1stTab} shows.
Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
because it is performed modulo $2$.
Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.

\begin{table}[bp]
% \vspace{-0.3cm}
\caption{Addition and multiplication operations over $\F_2$.\vspace{0.5cm}}
	\centering
	\label{tab:1stTab}
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
% 	\indent\\
	\hhline{~---~}
	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $1$ & \\
	& $1$ & $1$ & $0$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Addition over $\F_2$}\\
	\end{tabular}
	\quad
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
% 	\indent\\
	\hhline{~---~}
	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $0$ & \\
	& $1$ & $0$ & $1$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
	\end{tabular}

\end{table}

@@

\begin{table}[bp]
\begin{center}
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$.}\label{tab:gfelement} 
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
\caption{A typical SMSO structure of Massey-Omura multiplier.}
\label{fig:MasseyOmura}}
\end{figure}

@@

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
\caption{5-bit Agnew's SMPO. Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$. Index $w = 2i\pmod 5$.}
\label{fig:SMPO}}
\end{figure}

@@

@@

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{$A,B\in \Fkk$ given {\it w.r.t.} NB $N$}
 \KwOut{$R=A\times B$}
%%%%%%%%%%%%%%%%%%%%
  Initialize $A,B$ and aux var $X$ to 0\;
  \For { ($i=0$;   $i < k$; ++i ) }
  {
	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Equation \ref{eqn:aux}*/\;}
	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
  }
  $R\gets X$
\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
\end{algorithm}

In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
This is because of equation
$$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
sequence $F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
\end{Proof}

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
\caption{A 5-bit RH-SMPO.}
\label{fig:5bitRH}}
\end{figure}

To better understand the mechanism of RH-SMPO, we will use this 5-bit 
RH-SMPO as an example and introduce the details on how to design it.
\begin{Example}[Designing a 5-bit RH-SMPO]
From Equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
and XOR gates in $c$-layer according to Equation \ref{eqn:auxDC}. Concretely, as Algorithm \ref{alg:RHmulti}
describes, we implement auxiliary function $F_{k-1}$ in the logic:
$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
\begin{equation}
\label{eqn:5bitRHaux}
F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
\end{equation}
Consider equality of indices $4+1=0\bmod 5,~4+2=1\bmod 5$, we can write down gates in $c$-layer and $d$-layer (besides $d_0$)
$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
$i$-th gate forms the coefficient of $\beta^{2^i}$.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=3in]{newfig/multitable.pdf}
\caption{A $5\times 5$ multiplication table.}
\label{fig:multitable}}
\end{figure}

### 2017-01-24 dissertation/submission/prelim_basis.tex
dissertation after defense update 2
dissertation: after submission_1st revision

\begin{algorithm}[H]
\SetAlgoNoLine

 \KwIn{$f,f_{1},\dots,f_{s}$}
 \KwOut{$r,a_{1},\dots,a_{s}$, such that $f=a_{1}\cdot f_{1}+\dots+a_{s}\cdot f_{s}+r$.}

@@

\begin{Proof}
If $lt(f)$ and $lt(g)$ have no common variables,  
$L=lcm(lt(f),lt(g))=lt(f)\cdot lt(g)$. Then: 
\begin{equation}
    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g=
\frac{lt(f)\cdot lt(g)}{lt(f)}\cdot f - \frac{lt(f)\cdot lt(g)}{lt(g)}\cdot g
= lt(g)\cdot f - lt(f)\cdot g \nonumber
\end{equation}
Thus, every monomial in $Spoly(f, g)$ is divisible by either $lt(f)$ 
or $lt(g)$, so computing 
$Spoly(f, g) \stackrel{f,g}{\textstyle\longrightarrow}_+r$ will give $r=0$.
\end{Proof}

@@

\subsection{Elimination Theory}
Assume we are given a set of polynomials $f_1,\dots,f_s$ from a ring $\Fq[x_1,\dots,x_l,\dots,x_d]$.
First, we show that eliminating $x_1,\dots,x_l$ variables related to 
{\it projecting} the variety $V(\langle f_1,\dots,f_s\rangle)$ from 
$\Fq^d$ to $\Fq^{d-l}$. Figure \ref{fig:projection} is an example of projection in space of varieties
from $\Fq^3$ to $\Fq^2$, corresponding to eliminating variable $x_1$ from 
a system of polynomials belonging to $\Fq[x_1,x_2,x_3]$.

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/projection.pdf}
\caption{An example of projection from $\Fq^3$ to $\Fq^2$}
\label{fig:projection}}
\end{figure}

@@

Elimination theory provides the basis for following abstraction approach.
\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
\caption{Galois field arithmetic circuit model for abstraction}
\label{fig:abs_setup}}
\end{figure}

\begin{Problem}
\label{setup:abs}
Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of any circuit over $\Fq, q=2^k$, as Figure \ref{fig:abs_setup} shows.
In general, this circuit performs some unknown function $f$ over 
$\Fq$ such that every $f:\Fq\to\Fq$ is represented in polynomial form.
For any circuit $\mathcal C$, $Z = \Func(A)$. Then how to derive this function $\Func$
using GB?
\end{Problem}

The polynomial representation of $\Func$ over $\Fkk$ is:

\begin{equation}
f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
\end{equation}

@@

\begin{figure}[H]
\centerline{
\includegraphics[scale=0.5]{./figures/2bitmasmult.eps}
}
\caption{ A 2-bit multiplier over ${\mathbb{F}}_{2^2}$.}
\label{fig:2bitmul}
\end{figure}

>>>>

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{$f,f_{1},\dots,f_{s}$}
 \KwOut{$r,a_{1},\dots,a_{s}$, such that $f=a_{1}\cdot f_{1}+\dots+a_{s}\cdot f_{s}+r$.}

@@

\begin{Proof}
If $lt(f)$ and $lt(g)$ have no common variables,  
$L=lcm(lt(f),lt(g))=lt(f)\cdot lt(g)$. Then: 
\begin{equation}
    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g=
\frac{lt(f)\cdot lt(g)}{lt(f)}\cdot f - \frac{lt(f)\cdot lt(g)}{lt(g)}\cdot g
= lt(g)\cdot f - lt(f)\cdot g \nonumber
\end{equation}
Thus, every monomial in $Spoly(f, g)$ is divisible by either $lt(f)$ 
or $lt(g)$, so computing the remainder of \textit{Spoly}: 
$Spoly(f, g) \stackrel{f,g}{\textstyle\longrightarrow}_+r$ will give $r=0$.
\end{Proof}

@@

\subsection{Elimination Theory}
Assume we are given a set of polynomials $f_1,\dots,f_s$ from a ring $\Fq[x_1,\dots,x_l,\dots,x_d]$.
First, we show that eliminating $x_1,\dots,x_l$ variables related to 
{\it projecting} the variety $V(\langle f_1,\dots,f_s\rangle)$ from 
$\Fq^d$ to $\Fq^{d-l}$. Figure \ref{fig:3Dprojection} is an example of projection in space of varieties
from $\Fq^3$ to $\Fq^2$, corresponding to eliminating variable $x_1$ from 
a system of polynomials belonging to $\Fq[x_1,x_2,x_3]$.

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/projection.pdf}
\caption{An example of projection from $\Fq^3$ to $\Fq^2$.}
\label{fig:3Dprojection}}
\end{figure}

@@

Elimination theory provides the basis for following abstraction approach.
\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
\begin{Problem}
\label{setup:abs}
Let $S$ be the system of polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of any circuit over $\Fq, q=2^k$, as Figure \ref{fig:abs_setup} shows.
In general, this circuit performs some unknown function $f$ over 
$\Fq$ such that every $f:\Fq\to\Fq$ is represented in polynomial form.
For any circuit $\mathcal C$, $Z = \Func(A)$. Then how to derive this function $\Func$
using GB?
\end{Problem}

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
\caption{Galois field arithmetic circuit model for abstraction.}
\label{fig:abs_setup}}
\end{figure}

The polynomial representation of $\Func$ over $\Fkk$ is:
\begin{equation}
f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
\end{equation}

@@

\begin{figure}[bp]
\centerline{
\includegraphics[scale=0.5]{./figures/2bitmasmult.eps}
}
\caption{ A 2-bit multiplier over ${\mathbb{F}}_{2^2}$.}
\label{fig:2bitmul}
\end{figure}
